---
title: "First Draft: Singleton and Voids"
author: "Guna Pemmaraju"
date: "10/12/2020"
urlcolor: blue
linkcolor: blue
output:
  pdf_document:
    number_sections: yes
    toc: yes
    toc_depth: 2
---
\newpage

```{r setup, include=FALSE}
library(kableExtra)
source("./card_setup.R")
allDF <- read_csv("./output/SingletonVoids.csv")%>%
  mutate(VoidsInDeal=as.character(VoidsInDeal))%>%
  mutate(SingleTonsInDeal=as.character(SingleTonsInDeal))

singletonsDf <- allDF%>%group_by(SingleTonsInDeal)%>%
  summarise(n=sum(n), .groups="drop")%>%
  mutate(Percent=100*n/sum(n))%>%
  mutate(Percent=paste(round(Percent,2), "%"))

VoidsDf <- allDF%>%group_by(VoidsInDeal)%>%
  summarise(n=sum(n), .groups="drop")%>%
  mutate(Percent=100*n/sum(n))%>%
  mutate(Percent=paste(round(Percent,2), "%"))

knitr::opts_chunk$set(echo = FALSE)
```

# Introduction to the set-up

This simulation was setup using the standard deck of cards, and set to randomly generate deals. The source code used can be viewed at: https://github.com/gunawebs/guna-bridge-repo

Feel free to fork / send pull requests with bug fixes and improvements.




## A single simulated deal

A single simulated deal would look like this:


```{r Deal cards}
dealCards()%>%as.data.frame()
```




## Goals for the simulation effort

For the simulation, 1 million deals were generated. The first analysis, more as a sample case, consisted of looking for the distribution of singletons and voids. 
As is obvious, these could technically have been worked out using the basic probability theory. However:

- This Goal-Set is used as a way to showcase what can be done, especially in future, using the simulation approach
- Mathematical computation may not be easily explainable to everyone. On the other hand a simulation is very easy to understand. 
And with sufficiently large n, it gains user confidence with ease.

Some of the questions that the current simulation tried to answer are:

- What is the probability that deal has exactly one singleton?
- What is the probability that deal has at least one singleton?
- What is the probability that deal has exactly one void?
- What is the probability that deal has at least one void?
- What is the interplay of distribution between singletons and voids in a deal


\newpage


# Analysis of the outputs from Simulation

## Distribution of Singletons in a deal

```{r singletons, echo=FALSE}
singletonsDf%>%
  ggplot(aes(x=SingleTonsInDeal, y=n, label=Percent))+
  geom_label(vjust = -0.10)+
  geom_bar(stat="identity", fill="#445588")+
  labs(x="Singletons in a deal", y="Number of deals Per Million dealt")+
  scale_y_continuous(limits = c(0,400000), labels = comma)

```

As we can see, only 25% of deals are bereft of a singleton! Which means a staggering 75% of the deals would have at least 1 singleton. 

**Data Based Insights**

- If you do not see a singleton between you and the dummy, there is a 75% chance that one of the opponents has a singleton. So, do play for a bad break!
- If you are already seeing a singleton, the chance of one or more other singletons is around 51%
- 38% of the deals would have at least 2 singletons. So if the previous deal had less than 2 singletons, the chance of this deal having at least 2 singletons is *a whopping 76%!*
- 12% of the deals would have at least 3 singletons. So in an 8 board set, do not rule out a 3 singleton deal.

## Distribution of Voids in a deal

```{r voids, echo=FALSE}
VoidsDf%>%
  ggplot(aes(x=VoidsInDeal, y=n, label=Percent))+
  geom_label(vjust = -0.10)+
  geom_bar(stat="identity", fill="#445588")+
  labs(x="Voids in a deal", y="Number of deals Per Million dealt")+
  scale_y_continuous(limits = c(0,900000), labels = comma)

```

As we can see, about 18% of the deals have at least one void.

**Data Based Insights**

- In a 6 board match, at least one deal would have a void. 
- In a 12 board set, if you haven't seen any void in the first 5 boards, pretty much, any moment now a void will show up.

## Interplay between Voids and Singletons

```{r interplay, echo=FALSE}
allDF%>%
  ggplot(aes(x=SingleTonsInDeal, y=n, fill=VoidsInDeal))+
  geom_bar(position="stack", stat="identity")+
  geom_label(data=singletonsDf%>%mutate(VoidsInDeal="0"), mapping=aes(label=Percent), vjust = -0.10,  show.legend  = F)+
  labs(x="Singletons in a deal", y="Number of deals Per Million dealt")+
  scale_y_continuous(labels = comma, limits = c(0,400000))+scale_fill_manual(values=brewer.pal(8, "Accent"))

```

There seems to be an interesting interplay of voids and singletons and these needs to be looked at deeper.

## Distribution of Singletons in deals with no voids

```{r noVoidsSingletons, echo=FALSE}
allDF%>%
  filter(VoidsInDeal=="0")%>%group_by(SingleTonsInDeal)%>%
  summarise(n=sum(n), .groups="drop")%>%
  mutate(Percent=100*n/sum(n))%>%
  mutate(Percent=paste(round(Percent,2), "%"))%>%
  ggplot(aes(x=SingleTonsInDeal, y=n, label=Percent))+
  geom_label(vjust = -0.10)+
  geom_bar(stat="identity", fill="#445588")+
  labs(x="Singletons in a deal", y="Number of deals Per Million dealt")+
  scale_y_continuous(limits = c(0,400000), labels = comma)

```

There does not seem to be a big change in the probability distribution of this one.

## Distribution of Singletons in deals with 1 void

```{r oneVoidsSingletons, echo=FALSE}
allDF%>%
  filter(VoidsInDeal=="1")%>%group_by(SingleTonsInDeal)%>%
  summarise(n=sum(n), .groups="drop")%>%
  mutate(Percent=100*n/sum(n))%>%
  mutate(Percent=paste(round(Percent,2), "%"))%>%
  ggplot(aes(x=SingleTonsInDeal, y=n, label=Percent))+
  geom_label(vjust = -0.10)+
  geom_bar(stat="identity", fill="#445588")+
  labs(x="Singletons in a deal", y="Number of deals Per Million dealt")+
  scale_y_continuous(limits = c(0,80000), labels = comma)

```
**Data Based Insights**

- If you are seeing a void, probability that one of the others has a singleton is much higher than otherwise
\newpage

## Overall Cross distribution of Singletons Vs Voids

**Cross distribution of Singletons Vs Voids across Million Deals**

```{r overallTable, echo=FALSE}
tableRds <- readRDS("./output/rds_table_SingletonVoids.RDS")
kbl(tableRds)%>%
  kable_classic() %>%
  add_header_above(c("Singletons in the Deal" = 1, "Voids in the Deal" = 5))
```


**Converted to Probability Percentages**

```{r overallTablePercent, echo=FALSE}
res <- 100*tableRds/sum(tableRds)
myRes <- paste0(round(res,2), "%")
dat <- as.data.frame(matrix(myRes, nrow=nrow(res), ncol=ncol(res)))
colnames(dat) <-colnames(res)
rownames(dat) <-rownames(res)

kbl(dat)%>%
  kable_classic() %>%
  add_header_above(c("Singletons in the Deal" = 1, "Voids in the Deal" = 5))
```

## End Note
The above simulation and the subsequent analysis and this report was created in the last few hours, out a whim for simulation. However a solid platform has been set for further analysis. Apart from analyzing hand distribution, a double dummy engine could be plugged in, and with that a plethora of simulations, followed by insights, could be expected. 

PS: This is v1, and hence errors, including typos, are expected.


